Understanding the Impact of Socioeconomic and Regional Factors on Health and Educational Outcomes

### 2024-12-14 paper.qmd

| Variable                                 | **Inadequate School Funding**  | **Adequate School Funding**  |
|------------------------------------------|--------------------------------|------------------------------|
| Residential Segregation Index: Mean (SD) | 46 (17)                        | 56 (14)                      |
| Region-West: Count (%)                   | 114 (7.3%)                     | 89 (10.8%)                   |
| Region-Northeast: Count (%)              | 3 (0.2%)                       | 200 (24.3%)                  |
| Region-Midwest: Count (%)                | 267 (17.1%)                    | 285 (34.7%)                  |
| Region-South: Count (%)                  | 1173 (75.3%)                   | 248 (30.2%)                  |
| Median Property Tax, USD: Mean (SD)      | 1119 (763)                     | 2282 (1648)                  |
| Median Household Income, USD: Mean (SD)  | 56513 (13200)                  | 72218 (18255)                |

@@

@@

@@

| Statistic          | Residential Segregation Index (RSI) | Percentage With Limited Access to Healthy Foods | Percentage of Uninsured Adults | Median Household Income, USD | Median Property Tax, USD | Average Life Expectancy |
|--------------------|-------------------------------------|-------------------------------------------------|--------------------------------|------------------------------|--------------------------|-------------------------|
| Min                | 1.0                                 | 0.0                                             | 2.8                            | 28972                        | 0                        | 64.9                    |
| Median             | 50.3                                | 6.5                                             | 12.6                           | 58623                        | 1173                     | 75.1                    |
| Mean               | 49.3                                | 8.2                                             | 14.1                           | 61940                        | 1521                     | 75.2                    |
| Max                | 93.8                                | 62.6                                            | 47.1                           | 167605                       | 9788                     | 98.9                    |
| Standard Deviation | 16.5                                | 6.9                                             | 6.6                            | 16877.8                      | 1274.7                   | 3.1                     |

>>>>

| Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
|-----------------------------|----------------------|---------------------|
| Residential Segregation Index: Mean (SD) | 46 (17)                       | 56 (14)                     |
| Region-West: Count (%)                   | 114 (7.3%)                    | 89 (10.8%)                  |
| Region-Northeast: Count (%)              | 3 (0.2%)                      | 200 (24.3%)                 |
| Region-Midwest: Count (%)                | 267 (17.1%)                   | 285 (34.7%)                 |
| Region-South: Count (%)                  | 1173 (75.3%)                  | 248 (30.2%)                 |
| Median Property Tax, USD: Mean (SD)      | 1119 (763)                    | 2282 (1648)                 |
| Median Household Income, USD: Mean (SD)  | 56513 (13200)                 | 72218 (18255)               |

@@

```{r}
# Remove influential points from the dataset
combined_no_influential <- combined[-influential, ]

# Refit the model without influential points
mlr_no_influential <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac + X..Limited.Access.to.Healthy.Foods +X..Uninsured.Adults, data = combined_no_influential)

# Compare the two models
summary(mlr.imp)  # Original model
summary(mlr_no_influential)  # Model without influential points
```

```{r}
#looking for outliers 
library(car)
outlier_points <- outlierTest(mlr.imp)
print(outlier_points)
```

```{r}
outlier_indices <- as.numeric(rownames(outlier_points))  # Extract row indices
print(outlier_indices)  # View the indices to confirm

@@

```{r}
par(mfrow = c(2, 2), mar = c(4,4,2,2))
plot(mlr_no_influential)
```

@@

| Statistic          | Residential Segregation Index (RSI) | Percentage With Limited Access to Healthy Foods | Percentage of Uninsured Adults | Median Household Income, USD | Median Property Tax, USD | Average Life Expectancy |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
| Min                | 1.0                                 | 0.0                                             | 2.8                            | 28972                        | 0                        | 64.9                    |
| Median             | 50.3                                | 6.5                                             | 12.6                           | 58623                        | 1173                     | 75.1                    |
| Mean               | 49.3                                | 8.2                                             | 14.1                           | 61940                        | 1521                     | 75.2                    |
| Max                | 93.8                                | 62.6                                            | 47.1                           | 167605                       | 9788                     | 98.9                    |
| Standard Deviation | 16.5                                | 6.9                                             | 6.6                            | 16877.8                      | 1274.7                   | 3.1                     |
